---
title: "Math 180 Murder per Population Prediction"
author: "Jessie Arroyo"
date: "2022-12-05"
output: html_document
---

## Reading the Data and Renaming some of the column names
```{r Reading Data}
CrimeData <- read.csv("~/Downloads/CrimeData/crimedata.csv", na.strings="?", stringsAsFactors=TRUE)
colnames(CrimeData)[1] <- "Community Name" #Renaming
colnames(CrimeData)[3] <- "County Code" #Renaming
colnames(CrimeData)[4] <- "Community Code" #Renaming

# Viewing Dimension of the Dataset
dim(CrimeData)
```

## Using Min-Max standardization to normalize the data 
```{r Data Pre Processing}
library(caret)
pp = preProcess(CrimeData, method = "range")
crimedata = predict(pp, CrimeData)
crimedata = na.omit(crimedata)

#removing Community Name, State, Community Code 
crimedata <- crimedata[ -c(1,0:5) ]

#Viewing the amount of rows and columns
ncol(crimedata)
nrow(crimedata)

```


## Splitting the Data into training and testing. 
```{r}
set.seed(1)
train = sample(nrow(crimedata), 0.6*nrow(crimedata)) #splitting the data
CrimeTr = crimedata[train,] #Creating Training Data set
CrimeTe = crimedata[-train,] # Creating test Data 
dim(CrimeTr)
dim(CrimeTe)
```
## Lasso Regression: (Jessie)

```{r Murder Per Population}
x= model.matrix(murdPerPop ~., CrimeTr)[,-1]
y = CrimeTr$murdPerPop
xtest = model.matrix(murdPerPop ~., CrimeTe)[,-1]


set.seed(1)
library(glmnet)

# Fitting Model
out <- glmnet(x,y, alpha = 1)
plot(out)

#Cross Validation
cv.out = cv.glmnet(x, y, alpha = 1)
plot(cv.out)

#Predicting Model
lasso.pred = predict(cv.out, s = cv.out$lambda.min, newx = xtest)

# test MSE 
mean((lasso.pred - CrimeTe$murdPerPop)^2)

#Finding the Valuable Variables 
lasso.coef = predict(out,type = 'coefficients',s = cv.out$lambda.min)[1:142,]

# Variables that were Viewed to be important 
lasso.coef[lasso.coef != 0]
R2.murd = caret::R2(lasso.pred,CrimeTe$murdPerPop)
cat("\nR2: ", R2.murd, "\nTest MSE:", mean((lasso.pred - CrimeTe$murdPerPop)^2))
```

## Plotting Correlation
```{r Plotting correlation plot}
library(corrplot)
library(dplyr)

#Murders
murder_var <- crimedata %>% select(racepctblack, agePct65up, PctKidsBornNeverMar, PctRecentImmig, PersPerOwnOccHous, MedNumBR, PctVacantBoarded, MedYrHousBuilt, PctHousNoPhone, NumStreet, PctForeignBorn, LemasTotReqPerPop, PctPolicWhite, PolicAveOTWorked, LemasPctPolicOnPatr, PolicBudgPerPop, murdPerPop)
murd <- cor(murder_var)
corrplot(murd, method = 'color', order = 'alphabet', tl.cex = 0.5)
```

## Multilinear Regression (David)
```{r LM Murder per Population}
#Training data
model.murd <- lm(murdPerPop ~ racepctblack +  agePct65up+ PctKidsBornNeverMar+ PctRecentImmig+ PersPerOwnOccHous+ MedNumBR+ PctVacantBoarded+ MedYrHousBuilt+ PctHousNoPhone+ NumStreet+ PctForeignBorn+ LemasTotReqPerPop+ PctPolicWhite+ PolicAveOTWorked+ LemasPctPolicOnPatr+ PolicBudgPerPop,data = CrimeTr)
#Testing data
pred.murder = predict(model.murd, newdata = CrimeTe)
#Test MSE
mean((pred.murder - CrimeTe$murdPerPop)^2)

# Cross Validating
set.seed(1)
library(caret)
train.control <-trainControl(method = "cv", number = 10)
cv.model <- train(murdPerPop ~ racepctblack +  agePct65up+ PctKidsBornNeverMar+ PctRecentImmig+ PersPerOwnOccHous+ MedNumBR+ PctVacantBoarded+ MedYrHousBuilt+ PctHousNoPhone+ NumStreet+ PctForeignBorn+ LemasTotReqPerPop+ PctPolicWhite+ PolicAveOTWorked+ LemasPctPolicOnPatr+ PolicBudgPerPop,data = CrimeTr, method = "lm", trControl = train.control)
cv.model
summary(cv.model$finalModel)

#Test MSE
pred.cvmurd = predict(cv.model$finalModel, newdata = CrimeTe)
mean((pred.cvmurd-CrimeTe$murdPerPop)^2)
plot(cv.model$finalModel)
postResample(pred.cvmurd, CrimeTe$murdPerPop)
cat("Test MSE: ",mean((pred.cvmurd-CrimeTe$murdPerPop)^2))
```

#Removing Outlier to see if would improve
```{r}
library(caret)
remove.outlier<-CrimeTr[-1582,]
re <- CrimeTe[-1582,]

model <- lm(murdPerPop ~ racepctblack +  agePct65up+ PctKidsBornNeverMar+ PctRecentImmig+ PersPerOwnOccHous+ MedNumBR+ PctVacantBoarded+ MedYrHousBuilt+ PctHousNoPhone+ NumStreet+ PctForeignBorn+ LemasTotReqPerPop+ PctPolicWhite+ PolicAveOTWorked+ LemasPctPolicOnPatr+ PolicBudgPerPop,data = remove.outlier)
summary(model)
pred <- predict(model, re )
mean((pred-re$murdPerPop)^2)
postResample(pred, re$murdPerPop)

```
# Combining Highly Multicollinear variables into one 
```{r}
model.inter <- lm(murdPerPop ~ racepctblack+racepctblack*PctPolicWhite + racepctblack*PctVacantBoarded+ agePct65up+ PctKidsBornNeverMar+ +PctKidsBornNeverMar*NumStreet+ PctKidsBornNeverMar*PctVacantBoarded+ PctRecentImmig+ PersPerOwnOccHous+ MedNumBR+ PctVacantBoarded+ MedYrHousBuilt+ PctHousNoPhone+PctHousNoPhone*PctVacantBoarded+  NumStreet+ PctForeignBorn*PctRecentImmig+PctForeignBorn + LemasTotReqPerPop+ PctPolicWhite+ PolicAveOTWorked+ LemasPctPolicOnPatr+ PolicBudgPerPop,data =CrimeTr)
summary(model.inter)
pred.inter <- predict(model.inter, newdata = CrimeTe)
mean((pred.inter - CrimeTe$murdPerPop)^2)
postResample(pred.inter, CrimeTe$murdPerPop)
```

## K Nearest Neighbors (Oscar)
```{r KNN murder per population}
library(caret)
set.seed(1)
x_train = CrimeTr[c(3, 10, 51, 57, 71, 75, 77, 78, 93, 94, 104, 108, 115, 121, 124)]
y_train = CrimeTr$murdPerPop
x_test = CrimeTe[c(3, 10, 51, 57, 71, 75, 77, 78, 93, 94, 104, 108, 115, 121, 124)]
y_test = CrimeTe$murdPerPop

#CV the model 
cv.knnmurd <- train(murdPerPop ~ racepctblack +  agePct65up+ PctKidsBornNeverMar+ PctRecentImmig+ PersPerOwnOccHous+ MedNumBR+ PctVacantBoarded+ MedYrHousBuilt+ PctHousNoPhone+ NumStreet+ PctForeignBorn+ LemasTotReqPerPop+ PctPolicWhite+ PolicAveOTWorked+ LemasPctPolicOnPatr+ PolicBudgPerPop,data = CrimeTr, method = "knn", trControl = trainControl(method = "cv", number = 10), tuneLength = 10)
cv.knnmurd
plot(cv.knnmurd)

#Fitting With CV 
knn.murd <- knnreg(x_train, y_train, k = 5)
pred_ymurd = predict(knn.murd, x_test)

mse = mean((pred_ymurd- y_test)^2)
postResample(pred_ymurd, CrimeTe$murdPerPop)
cat("MSE: ", mse)

x = 1:length(y_test)
plot(x, y_test, col = "red", type = "l", lwd=2,
     main = "Murder per Population Prediction", xlab = "Sample Size", ylab = "Actual Murder per Population")
lines(x, pred_ymurd, col = "blue", lwd=2)
legend("topright",  legend = c("Actual-Murder per Population", "predicted-Murder per Population"), 
       fill = c("red", "blue"), col = 2:3,  adj = c(0, 0.6))
grid() 
```


## Random Forest (Rebecca)
```{r RF Murders Per Population}
#new set of data only having the predictors that LASSO said were important
murdPredictors = crimedata[c(3, 10, 51, 57, 71, 75, 77, 78, 93, 94, 104, 108, 115, 121, 124, 126)]
#Creating the trees
library(randomForest)
set.seed(1)

murder.crime.rf.train <- randomForest( murdPerPop ~ ., data = murdPredictors, subset = train, mtry = 12, importance = TRUE)
yhat.murder <- predict(murder.crime.rf.train, newdata = crimedata[-train, ])
mean((yhat.murder - CrimeTe$murdPerPop)^2)
#Seeing what the importance of each of the predictors were
importance(murder.crime.rf.train)
varImpPlot(murder.crime.rf.train, cex = 0.5)
murder.crime.rf.train
library(caret)
postResample(yhat.murder, CrimeTe$murdPerPop)
cat("Test MSE: ",mean((yhat.murder - CrimeTe$murdPerPop)^2))

```


## Regression Trees(Jessie)
```{r RT murder per population}
set.seed(1)
library(tree)
tree.murder <-tree(murdPerPop ~ ., data = murdPredictors, subset = train)
summary(tree.murder)
plot(tree.murder)
text(tree.murder, pretty = 0)
title(main = "Unprunned Tree")

#Checking if prunning will improve peformance
cv.murder <- cv.tree(tree.murder, K = 10)
plot(cv.murder$size, (cv.murder$dev), type = "b", xlab = "Tree Size", ylab = "CV-Error Rate")

#Prunning the tree
prune.murd <- prune.tree(tree.murder, best = 4)
plot(prune.murd)
text(prune.murd, pretty = 0)
title("Prunned Pree")

# Test Set MSE using Unpruned Tree
yhat.treemurd <- predict(tree.murder, newdata = CrimeTe)
plot(yhat.treemurd, CrimeTe$murdPerPop, xlab = "Predicted Unprunned Murders per Population", ylab = "Actual Murders per Population")
abline(0,1)

unprunemse = mean((yhat.treemurd - CrimeTe$murdPerPop)^2)

postResample(yhat.treemurd, CrimeTe$murdPerPop)

cat("Unprunned Prunned Test MSE: ", unprunemse)

# TEST MSE for Prunned Tree
yhat.prunemurd <- predict(prune.murd, newdata = CrimeTe)
prunemse = mean((yhat.prunemurd - CrimeTe$murdPerPop)^2)
plot(yhat.prunemurd, CrimeTe$murdPerPop, xlab = "Predicted Prunned Murders per Population", ylab = "Actual Murders per Population")
abline(0,1)

postResample(yhat.prunemurd, CrimeTe$murdPerPop) #Prunned
cat("Prunned Test MSE: ", prunemse)
```