---
title: "Math 180 Project"
author: "Jessie Arroyo, Rebecca Serrato, David Lee, Oscar Cortes Torres"
date: "2022-09-21"
output: html_document
---


## Reading the Data and Renaming some of the column names
```{r}
#setwd("~/Downloads/CrimeData") Setting a Directory
CrimeData <- read.csv("~/Downloads/CrimeData/crimedata.csv", na.strings="?", stringsAsFactors=TRUE)
colnames(CrimeData)[1] <- "Community Name" #Renaming
colnames(CrimeData)[3] <- "County Code" #Renaming
colnames(CrimeData)[4] <- "Community Code" #Renaming

# Viewing Dimension of the Dataset
dim(CrimeData)
```

## Using Min-Max standardization to normalize the data 

1. Download caret, to download caret, we click on "tools" then after click on "install packages", then search for "carnet" and click install 
```{r}
#minMax <- function(x) {
   # (x - min(x)) / (max(x) - min(x))
 # }
#normalizeCrimeData <- as.data.frame(lapply(CrimeData[6:147], minMax))
#apply Min-Max normalization to first four columns in iris dataset
#iris_norm <- as.data.frame(lapply(iris[1:4], min_max_norm))
#normalizeCrimeData

library(caret)
pp = preProcess(CrimeData, method = "range")
crimedata = predict(pp, CrimeData)
#Viewing Data
crimedata = na.omit(crimedata)

#removing Community Name, State, Community Code 
crimedata <- crimedata[ -c(1,0:5) ]

#Viewing the amount of rows and columns
ncol(crimedata)
nrow(crimedata)

```

## Splitting the Data into training and testing. 
```{r}
train = sample(nrow(crimedata), 0.6*nrow(crimedata)) #splitting the data
CrimeTr = crimedata[train,] #Creating Training Data set
CrimeTe = crimedata[-train,] # Creating test Data 
dim(CrimeTr)
dim(CrimeTe)
```

## Lasso Regression:Viewing The most imporant variables for Each Response Variable. 
1. Murder per Population
2. Robberies per Population
3. Rapes per Population
4. Assaults per Population
5. Larcenies per Population
6. Auto Thief per population
7. Arson per Population

1. Murder Per Population- Lasso
```{r}
x= model.matrix(murdPerPop ~., CrimeTr)[,-1]
y = CrimeTr$murdPerPop
xtest = model.matrix(murdPerPop ~., CrimeTe)[,-1]
xcrime = model.matrix(murdPerPop ~., crimedata)[,-1]
ycrime = crimedata$murdPerPop
set.seed(1)

library(glmnet)
#Cross Validation
cv.out = cv.glmnet(x, y, alpha = 1)

#Predicting Model
lasso.pred = predict(cv.out, s = cv.out$lambda.min, newx = xtest)
mean((lasso.pred - CrimeTe$murdPerPop)^2)

#Fitting Model
out = glmnet(xcrime,ycrime, alpha = 1)

#Finding the Valuable Variables 
lasso.coef = predict(out,type = 'coefficients',s = cv.out$lambda.min)[1:142,]
```

2. Robberies Per Population
```{r}
xrobb= model.matrix(robbbPerPop ~., CrimeTr)[,-1]
yrobb = CrimeTr$robbbPerPop
xtestrobb = model.matrix(robbbPerPop ~., CrimeTe)[,-1]
xcrimerobb = model.matrix(robbbPerPop ~., crimedata)[,-1]
ycrimerobb = crimedata$robbbPerPop
set.seed(1)

library(glmnet)
#Cross Validation
cv.outrobb = cv.glmnet(xrobb, yrobb, alpha = 1)

#Predicting Model
lasso.predrobb = predict(cv.outrobb, s = cv.outrobb$lambda.min, newx = xtest)
mean((lasso.predrobb - CrimeTe$robbbPerPop)^2)

#Fitting Model
robb.out = glmnet(xcrimerobb,ycrimerobb, alpha = 1)

#Finding the Valuable Variables 
lasso.coefrobb = predict(robb.out,type = 'coefficients',s = cv.outrobb$lambda.min)[1:142,]
```

3. Rapes per Population(rpp)
```{r}
x_rpp= model.matrix(rapesPerPop ~., CrimeTr)[,-1]
y_rpp = CrimeTr$rapesPerPop
xtest_rpp = model.matrix(rapesPerPop ~., CrimeTe)[,-1]
xcrime_rpp = model.matrix(rapesPerPop ~., crimedata)[,-1]
ycrime_rpp = crimedata$rapesPerPop
set.seed(1)

library(glmnet)
#Cross Validation
cv.outrpp = cv.glmnet(x_rpp, y_rpp, alpha = 1)

#Predicting Model
lasso.predrpp = predict(cv.outrpp, s = cv.outrpp$lambda.min, newx = xtest_rpp)
mean_rpp =mean((lasso.predrpp - CrimeTe$rapesPerPop)^2)

#Fitting Model
fit_rpp = glmnet(xcrime,ycrime, alpha = 1)

#Finding the Valuable Variables 
lasso.coef_rpp = predict(fit_rpp,type = 'coefficients',s = cv.outrpp$lambda.min)[1:142,]

```

4. Assaults per Population(app)

```{r}
x_app= model.matrix(assaultPerPop ~., CrimeTr)[,-1]
y_app = CrimeTr$assaultPerPop 
xtest_app = model.matrix(assaultPerPop  ~., CrimeTe)[,-1]
xcrime_app = model.matrix(assaultPerPop  ~., crimedata)[,-1]
ycrime_app = crimedata$assaultPerPop 
set.seed(1)

library(glmnet)
#Cross Validation
cv.outapp = cv.glmnet(x_app, y_app, alpha = 1)

#Predicting Model
lasso.predapp = predict(cv.outapp, s = cv.outapp$lambda.min, newx = xtest_app)

#Test MSE
mean((lasso.predapp - CrimeTe$burglPerPop)^2)

#Fitting Model
fit_app = glmnet(xcrime_app,ycrime_app, alpha = 1)

#Finding the Valuable Variables 
lasso.coefapp = predict(fit_app,type = 'coefficients',s = cv.outapp$lambda.min)[1:142,]

```

5. Larcenies per Population(lpp)
```{r}
xlpp= model.matrix(larcPerPop ~., CrimeTr)[,-1]
ylpp = CrimeTr$larcPerPop
xtest_lpp = model.matrix(larcPerPop ~., CrimeTe)[,-1]
xcrime_lpp = model.matrix(larcPerPop ~., crimedata)[,-1]
ycrime_lpp = crimedata$larcPerPop
set.seed(1)

library(glmnet)
#Cross Validation
cv.out_lpp = cv.glmnet(xlpp, ylpp, alpha = 1)

#Predicting Model
lasso.pred_lpp = predict(cv.out_lpp, s = cv.out_lpp$lambda.min, newx = xtest_lpp)
mean((lasso.pred_lpp - CrimeTe$larcPerPop)^2)

#Fitting Model
lpp.out = glmnet(xcrime_lpp,ycrime_lpp, alpha = 1)

#Finding the Valuable Variables 
lasso.coeflpp = predict(lpp.out,type = 'coefficients',s = cv.out_lpp$lambda.min)[1:142,]

```

6. Auto Theif per population(atpp)
```{r}
xatpp= model.matrix(autoTheftPerPop ~., CrimeTr)[,-1]
yatpp = CrimeTr$autoTheftPerPop
xtest.atpp = model.matrix(autoTheftPerPop ~., CrimeTe)[,-1]
xcrime.atpp = model.matrix(autoTheftPerPop ~., crimedata)[,-1]
ycrime.atpp = crimedata$autoTheftPerPop
set.seed(1)

library(glmnet)
#Cross Validation
cv.outatpp = cv.glmnet(xatpp, yatpp, alpha = 1)

#Predicting Model
lasso.predatpp = predict(cv.outatpp, s = cv.outatpp$lambda.min, newx = xtest.atpp)

#test MSE
mean((lasso.predatpp - CrimeTe$autoTheftPerPop)^2)

#Fitting Model
atpp.out = glmnet(xcrime.atpp,ycrime.atpp, alpha = 1)

#Finding the Valuable Variables 
lasso.coefatpp = predict(atpp.out,type = 'coefficients',s = cv.outatpp$lambda.min)[1:142,]
```

7. Arson per Population(arpp)
```{r}
xarpp= model.matrix(arsonsPerPop ~., CrimeTr)[,-1]
yarpp = CrimeTr$arsonsPerPop
xtest.arpp = model.matrix(arsonsPerPop ~., CrimeTe)[,-1]
xcrime.arpp = model.matrix(arsonsPerPop ~., crimedata)[,-1]
ycrime.arpp = crimedata$arsonsPerPop
set.seed(1)

library(glmnet)
#Cross Validation
cv.outarpp = cv.glmnet(xarpp, yarpp, alpha = 1)

#Predicting Model
lasso.predarpp = predict(cv.outarpp, s = cv.outarpp$lambda.min, newx = xtest.arpp)

#Test MSE 
mean((lasso.predarpp - CrimeTe$arsonsPerPop)^2)

#Fitting Model
arpp.out = glmnet(xcrime.arpp,ycrime.arpp, alpha = 1)

#Finding the Valuable Variables 
lasso.coefarpp = predict(arpp.out,type = 'coefficients',s = cv.outarpp$lambda.min)[1:142,]
```

```{r}
library(e1071)
modelsvm = svm(murdPerPop ~racepctblack + PctRecentImmig + MedNumBR + PctVacantBoarded + PctHousNoPhone + NumStreet + PctPolicWhite + PctPolicBlack + PctPolicAsian + PctPolicMinor + PopDens + LemasPctPolicOnPatr + PolicBudgPerPop, CrimeTr)
pred.svm = predict(modelsvm, newdata = CrimeTe)

points(CrimeTe$racepctblack, )
```





